Formal Languages and Automatas

### 2024-08-16 paper/main.tex
done

\vspace*{10mm}
\begin{center}
\includegraphics[height=40mm]{sigillum.png}
\end{center}
\vspace*{10mm}

\title{Formal Languages and Automatas}
\date{\vspace{-3ex}}
{\let\newpage\relax\maketitle}
\thispagestyle{empty}

\begin{center}
\begin{large}
\begin{Large}
Seminar Paper on Formal Languages and Automatas\\
\end{Large}
for the course ``Algebra \& Computer Science'' \\
\end{large}
\end{center}
\vspace{1cm}
\begin{center}
\begin{large}
Supervisor: Prof. Dr. Jasmin Blanchette\\
\end{large}
\end{center}
\begin{center}
\begin{large}
Advisor: Xavier Genereux\\
\end{large}
\end{center}

@@

\section*{Abstract}
Abstract...
\textbf{Keywords}: Math

@@

\section{Introduction}

@@

First of all, we need to define some structures and ideas,
that will be really important in this paper. We will define 
formal languages in two ways and then automatas arcordingly.

@@

a series of characters from an alphabet.

@@

can really be anything. We could also define an Alphabet as subset of
the natural numbers $\Sigma \subset \mathbb{N}$. This is equalivant to
the previous definition, becaus we can always find a one-to-one map between

@@

Furhtermore its unintuitive to have words full of numbers and languages
full of number series. That is why im chosing this definition, like most

@@

Now with our first building block defined we can define words, which are just

@@

The empty word $\epsilon$ is analogous to the empty set in Set theory.
It has a lot uf uses, but one obvious is that now one can define 
a Monoid over words $(M, \cdot, \Sigma, \epsilon)$.
M is a set of words and multiplication is concetanation of words, then 
$\epsilon$ is the neutral element of the Monoid. Which brings us to

@@

In simple terms a concatenation of two words is just a new word with 
the two words chained togehter.\\
Now we can again build on words to define langauges.

@@

define language shortly. Obviously the simpler one has some downsides, 

@@

The kleene star $^*$ is a really powerful operator, to define it we need
to first define contenation for languages.

@@

In simple terms we are creating a new language by compbining every word
of the first language with every word of the second.
Now we are able to define the kleene star operator.

@@

@@

In simple terms it generates all possible permutation of abriatary length
of a formal language. This is really powerful, because $\Sigma^*$ is the
biggest language one can construct from the alphabet $\Sigma$. Thus every 
language $L$ generated by $\Sigma$ is just a subset of $\Sigma^*$.\\
Unintuively, we can also write a formal language as a formal power series.

@@

@@

This looks like a sum, we call it series, but it is actually a map from 
words to zero or one. The sum is a way to write a function definition: Infront
of every word $w \in \Sigma^*$ in the series, there is a $x \in \mathbb {B}$,
which is zero or one. So we can interpret every element in the series as an 

@@

All words that map to one, are in the language, the others are not. So we can write a formal language
as a function $\Sigma^* \to \mathbb{B}$.\\
The representation with sets has the andvantage that a formal language

@@

Those operations are really useful, thus we will define similar on the power

@@

@@

In simple terms, an finite automata gets an input string and a starting 
state. Then it works through the input string, starting at the first charater,
carrying the starting state to another according to the transition function.
Exactly the same procedure, for the second input string character and the 
remainding characters. If this process ends up in an accepting state, 
the automata accepts the input string. Consequently, 
the finite automata generates a language, namely the words that the automata 
accepts. The definition above is one possible way to encode a finite automata. 
Like with graphs we can encode the transition function of an finite

@@

  under matrix multiplication, nor matrix addition.\\

@@

@@

  Now we can write our finite automata in a shorter way:

@@

the state set, since as mentioned above, there always exist a one-to-one map
between a set with finite objects and a subset of the natural numbers. The first
object gets the number one, the second the number two, and so forth. The new
definition of the state transition function works like a lookup table, when one
wants to know what characters map state $i$ to state $j$, its a simple lookup:

@@

In this chapter we will discuss the obvious diffrences and 
simillarities of both representations. First we will give intuition,
why one might use a seires to represent a function definition.

@@

are really close to a series. For addition, formal power series and 
normal series are quite simillar, namely defined pointwise.\\

@@

to the cauchy product of series.

@@

Consequently, it is quite handy to use a formal power series as function 
definition, because the operations seem natural with series.

@@

%Maybe subsection
Next we will show that bot representations of a formal language are 
equivalent, by mapping one to another.
To be sure that both representation are actually equivalent, we need to
show there exist an bijective map, between formal power series and the
set of words. This map needs to preserve the structure, that means the
order of function application and operator application must be irrelevant.
Such maps are called isomporphism and structures for that a isomorphism 
exist are called isomporphic to each other. Thus we want tho show that
both representation are isomorphic.

@@

Now we have shown that both definition are equivalent, but why should
one use the less natural formal power series definition? The anwser
to this, as so often in mathematics, is {\bf abstraction}. Finite automata
are accepting formal languages, thus we can say they generate a formal
language. When abstracting from finite automata's to weighted automata's
which have a weight or cost between each transition, the generated
object is not longer a language, but a map that assigns a weight or
cost to each word. Generally, this weight or cost can be seen as a 
semiring (a Ring that does not have necessarily an additive inverse),
which are objects that behave like numbers, we will denote an arbitrary
semiring with $\mathbb{S}$. Those maps, that get accepted by an weighted

@@

indeed a formal power series over $\mathbb{S}$, consequently $L \in \fps{S}$.
To Summarize, finite automata generate formal power series over $\mathbb{B}$
and weighted automata generate formal power series over $\mathbb{S}$. So 
formal power series are in fact an abstraction of formal languages.

@@

In this chapter we will examine, what kind of languages get accepted 
by an finite automata. In order to do that, we will prove that all
languages that get accepted by an finite automata are rational and every
rational language can be generated by an finite automata (kleene's theorem).
A rational language is defined inductively.

@@

So we start with just one character and then building with our operations
step by step  a language. To show kleene's theorem we have to prove two
direction. First we will prove that finite automata only accept rational 
languages. Afterwards, we will prove that every rational language is 
accepted by an finite automata. In an effort to do that, we need to 
understand certain properties of the Matrix encoding of finite automatas.

@@

As mentioned before, this matrix is a transistion function of an finite
automata, where we can look up, what character carry one state in another.
In $A_{ij}$ is the character that carry $i$ to $j$, thus multiplying
$A$ by itself, gives us a new Matrix with all words with length two,
given as input and $i$ is the starting node would
carry $i$ to $j$ in two steps. In the formula above we see that we
are trying to go throug all nodes $k \in \{1, 2, 3, \dots n\}$

@@

This is not beatiful at all, but the idea stays exactly the same: Now
we have in $(A^n)_{ij}$ all words that carry $i$ to $j$ in n steps,
brute forcing throug all possible paths. Now to get all possible paths
of arbitrary length we just add all together:

@@

Yielding, that in $(A^*)_{ij}$ are all possible words, carrying
the automata from $i$ to $j$. If $i$ is the starting node and $j$
the accepting node, the accepting language of the automata is just
$(A^*)_{ij}$. That means, $A^*$ stores all accepting languages of 
an automata, with transition matrix $A$.

@@

@@

  arbitrary finite automata. $A$ is rational, because the transition

@@

  language of that finite autotamata, for every possible initial

@@

  We know that $A \in \fps{B}$ is rational, but that also means
  that $A^*$ is per definition rational, which finishes the base case.\\
  {\bf Inudction step: $n \to n+1$}\\

@@

also holds for $n+1$. we can assume that a transition matrix with dimenion 
$n \times n$ has only rational entries. \\

@@

Our goal is it to reduce the matrix $A$ by one dimension. In order to do
that we devide the matrix in four blocks:

@@

thus having two states. Our new two states consists of the original states

@@

\section{Maximum Gap Lemma}
\section{Not all Formal languages are regular}
\section{Kleene Schützenberg theorem and weigthed automatas}
\section{Conclusion}

\subsection{Citation} 
\label{sec:cit}

The citation method follows the author-year system. Place reference is in the text, footnotes should only be used for explanations and comments. The following notes are taken from the \emph{language} bibliography template from \url{ron.artstein.org}:\newline

\noindent
The \emph{Language} style sheet makes a distinction between two kinds of in-text citations: citing a work and citing an author.
\begin{itemize}
\item Citing a work:
  \begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parsep}{0pt}
  \item Two authors are joined by an ampersand (\&).
  \item More than two authors are abbreviated with \emph{et al.}
  \item No parentheses are placed around the year (though parentheses
    may contain the whole citation). 
  \end{itemize}
\item Citing an author:
  \begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parsep}{0pt}
  \item Two authors are joined by \emph{and}.
  \item More than two authors are abbreviated with \emph{and colleagues}.
  \item The year is surrounded by parentheses (with page numbers, if
    present).
  \end{itemize} 
\end{itemize}
To provide for both kinds of citations, \verb+language.bst+ capitalizes on the fact that \verb+natbib+ citation commands come in
two flavors. In a typical style compatible with \verb+natbib+, ordinary commands such as \verb+\citet+ and \verb+\citep+ produce short
citations abbreviated with \emph{et al.}, whereas starred commands such as \verb+\citet*+ and \verb+\citep*+ produce a citation with a
full author list. Since \emph{Language} does not require citations with full authors, the style \verb+language.bst+ repurposes the starred commands to be used for citing the author. The following table shows how the \verb+natbib+ citation commands work with \verb+language.bst+.
\begin{center}
  \begin{tabular}{lll}
    \toprule
    Command & Two authors & More than two authors \\
    \midrule
    \verb+\citet+ & \citet{hale} & \citet{sprouse} \\
    \verb+\citet*+ & \citet*{hale} & \citet*{sprouse} \\
    \addlinespace
    \verb+\citep+ & \citep{hale} & \citep{sprouse} \\
    \verb+\citep*+ & \citep*{hale} & \citep*{sprouse} \\
    \addlinespace
    \verb+\citealt+ & \citealt{hale} & \citealt{sprouse} \\
    \verb+\citealt*+ & \citealt*{hale} & \citealt*{sprouse} \\
    \addlinespace
    \verb+\citealp+ & \citealp{hale} & \citealp{sprouse} \\
    \verb+\citealp*+ & \citealp*{hale} & \citealp*{sprouse} \\
    \addlinespace
    \verb+\citeauthor+ & \citeauthor{hale} & \citeauthor{sprouse} \\
    \verb+\citeauthor*+ & \citeauthor*{hale} & \citeauthor*{sprouse} \\
    \verb+\citefullauthor+ & \citefullauthor{hale} & \citefullauthor{sprouse} \\
    \bottomrule
  \end{tabular}
\end{center}
Authors of \emph{Language} articles would typically use \verb+\citet*+, \verb+\citep+, \verb+\citealt+ and \verb+\citeauthor*+, though they
could use any of the above commands. There is no command for giving a full list of authors.

\subsection{Bibliography}
The bibliography of this template includes the references of the \emph{language} stylesheet as a sample bibliography.

@@

\microtypesetup{protrusion=false}
\listoffigures{}
\listoftables{}
\microtypesetup{protrusion=true}

\clearpage
\printglossaries
\pagebreak
\addcontentsline{toc}{section}{Bibliography}
\pagestyle{fancy}
\bibliographystyle{language-dt} 
\bibliography{bibliography} 
\nocite{*}
\end{document}

>>>>

\vspace*{10mm}
\begin{center}
\includegraphics[height=40mm]{sigillum.png}
\end{center}
\vspace*{10mm}
\title{Formal Languages and Automatas}
\date{\vspace{-3ex}}
{\let\newpage\relax\maketitle}
\thispagestyle{empty}

\begin{center}
\begin{large}
\begin{Large}
Seminar Paper on Formal Languages and Automata\\
\end{Large}
for the course ``Algebra \& Computer Science'' \\
\end{large}
\end{center}
\vspace{1cm}
\begin{center}
\begin{large}
Supervisor: Prof. Dr. Jasmin Blanchette\\
\end{large}
\end{center}
\begin{center}
\begin{large}
Advisor: Xavier Genereux\\
\end{large}
\end{center}

@@

@@

\section{Introduction}
This paper aims to give a gentle introduction to the world of
automata with formal power series. Formal power series are really
useful for weighted automata, but not so much for the topics we are
covering in this paper. Thus, some proofs may not be as simple as in 
other literature. First, we will go over some definitions, starting
from scratch, but ending up with the definition of an automaton using
a matrix and  formal power series. In the third chapter, we will
provide some intuition as to why one would use a formal power
series in this context and show that the definition of formal 
language using formal power series and sets is equivalent.
At the end of the third chapter, we will discuss why formal power 
series are so powerful in the context of automata theory.
In the fourth chapter, we will prove the kleene's theorem using the
formal power series representation of sets. 
At the end, we will give  an outlook on weighted automata.

@@

Firstly, we need to define some structures and ideas, 
that will be really important in this paper.
We will define  formal languages in two ways and then 
the automata accordingly. 

@@

a series of characters from an alphabet

@@

can really be anything. We could also define an alphabet as a subset of
the natural numbers $\Sigma \subset \mathbb{N}$. This is equivalent to
the previous definition, because we can always find a one-to-one map between

@@

Furthermore, it's unintuitive to have words full of numbers and languages full of number series. That is why we are choosing this definition, like most

@@

Now that our first building block is defined, we can define words, which are just

@@

The empty word $\epsilon$ is analogous to the empty set in set theory.
It has a lot of uses, but one obvious one is that now one can define 
a monoid over words $(M, \cdot, \Sigma, \epsilon)$.
M is a set of words and multiplication is the concatenation of words, then 
$\epsilon$ is the neutral element of the monoid. Which brings us to

@@

In simple terms, a concatenation of two words is just a new word with
the two words were chained together. Now we can again build on
words to define languages.

@@

define language shortly. Obviously, the simpler one has some downsides, 

@@

The kleene star $^*$ is a really powerful operator, to define it, we need
to first define contention for languages.

@@

In simple terms, we are creating a new language by
combining every word of the first language with every word of
the second. Now we are able to define the kleene star operator

@@

    L\hat{\mkern6mu} := \bigcup_{i \in \mathbb{N}} L^i = 
     L^1 \cup L^2 \cup L^3 \cup \dots

@@

In simple terms, it generates all possible permutations of arbitrary length of a formal language. This is really powerful, because $\Sigma^*$ is the
biggest language one can construct from the alphabet $\Sigma$. Thus, every  language $L$ generated by $\Sigma$ is just a subset of $\Sigma^*$.\\
Unintuitively, we can also write a formal language as a formal power series.

@@

  The language with no words is denoted by $L = 0$.

@@

This looks like a sum, we call it a series, but it is actually a map from 
words to zero or one. The sum is a way to write a function definition: In front of every word $w \in \Sigma^*$ in the series, there is a $x \in \mathbb {B}$, which is zero or one. So we can interpret every element in the series as an 

@@

All words that map to one are in the language, the others are not. So we can write a formal language as a function $\Sigma^* \to \mathbb{B}$.\\
The representation with sets has the advantage that a formal language

@@

Those operations are really useful, so we will define similar ones on the power

@@

    L\hat{\mkern6mu} :=  L^1 + L^2 + L^3 + \dots = \sum_{i \in \mathbb{N}_0} L^i.

@@

In simple terms, a finite automaton gets an input string and a starting  state. Then it works through the input string, starting at the first character and carrying the starting state to another according to the transition function.
Exactly the same procedure applies for the second input string character and the remaining characters. If this process ends up in an accepting state, the automata accept the input string. Consequently, the finite automata generate a language, namely the words that the automata 
accepts. The definition above is one possible way to encode a finite automaton. 
Like with graphs, we can encode the transition function of a finite

@@

  under matrix multiplication, nor matrix addition.
  \pagebreak\\

@@

    A\hat{\mkern6mu} = A + A^2 + A^3 \dots  = \sum_{i \in \mathbb{N}} A^i

@@

  Now, we can write our finite automata in a shorter way:

@@

the state set, since, as mentioned above, there always exists a one-to-one map between a set with finite objects and a subset of the natural numbers. The first object gets the number one, the second the number two, and so forth. The new definition of the state transition function works like a lookup table, when one wants to know what characters map state $i$ to state $j$, it is a simple lookup:

@@

In this chapter, we will discuss the obvious differences and  similarities between both representations. First, we will give an intuition as to why one might use a series to represent a function definition.

@@

are really close to a series. In addition, formal power series and 
Normal series are quite similar, namely defined pointwise.\\

@@

to the Cauchy product of series.

@@

Consequently, it is quite handy to use a formal power series as a function definition, because the operations seem natural with series.

@@

Next, we will show that both representations of a formal language are  equivalent, by mapping one to another.
To be sure that both representations are actually equivalent, we need to show there exists a bijective map between the formal power series and the set of words. 
This map needs to preserve the structure, which means the order of function application and operator application must be irrelevant.
Such maps are called isomorphisms, and the structures for that an isomorphism 
exist are called isomorphic to each other. Thus, we want to show that both representation are isomorphic.

@@

Now we have shown that both definitions are equivalent, but why should one use the less natural formal power series definition? The answer to this, as so often in mathematics, is {\bf abstraction}. Finite automata are accepting formal languages, thus we can say they generate a formal language.
 When abstracting from finite automata's to weighted automata's, which have a weight or cost between each transition, the generated object is not longer a language, but a map that assigns a weight or cost to each word. 
Generally, this weight or cost can be seen as a semi-ring (a ring that does not necessarily have an additive inverse), which are objects that behave like numbers. We will denote an arbitrary
semi-ring with $\mathbb{S}$. Those maps, that get accepted by a weighted

@@

indeed, a formal power series over $\mathbb{S}$, consequently $L \in \fps{S}$.
To summarize, finite automata generate formal power series over $\mathbb{B}$
and weighted automata generate formal power series over $\mathbb{S}$. So  formal power series are, in fact, an abstraction of formal languages.

@@

In this chapter, we will examine, what kinds of languages get accepted by a finite automaton. In order to do that, we will prove that all languages that get accepted by an finite automaton are rational, and every rational language can be generated by a finite automaton (kleene's theorem).
A rational language is defined inductively.\\

@@

So we start with just one character and then building with our operations step by step  a language. 
To show kleene's theorem, we have to prove two directions. 
First, we will prove that finite automata only accept rational  languages. 
Afterward, we will prove that every rational language is accepted by a finite automaton. 
In an effort to do that, we need to understand certain properties of the matrix encoding of finite automata.

@@

As mentioned before, this matrix is a transition function of a finite
automata, where we can look up what character carries one state in another.
In $A_{ij}$ is the character that carries $i$ to $j$, thus multiplying
$A$ by itself gives us a new matrix with all words of length two, given as input, and $i$ is the starting node that would carry $i$ to $j$ in two steps. In the formula above, we see that we
are trying to go through all nodes $k \in \{1, 2, 3, \dots n\}$

@@

This is not beautiful at all, but the idea stays exactly the same: now we have in $(A^n)_{ij}$ all words that carry $i$ to $j$ in n steps, brute forcing through all possible paths. Now to get all possible paths of arbitrary length, we just add them all together:

@@

Yielding, that in $(A^*)_{ij}$ are all possible words, carrying the automata from $i$ to $j$. If $i$ is the starting node and $j$ the accepting node, the accepting language of the automata is just
$(A^*)_{ij}$. That means, $A^*$ stores all accepting languages of  an automaton, with transition matrix $A$.

@@

\\ 
\\

@@

  arbitrary finite automata. $A$ is rational because the transition

@@

  language of that finite automata, for every possible initial

@@

We know that $A \in \fps{B}$ is rational, but that also means  that $A^*$ is per definition rational, which finishes the base case.\\
{\bf Induction step: $n \to n+1$}\\

@@

also holds for $n+1$. Thus, we can assume that a transition matrix with
dimension $n \times n$ has only rational entries. \\

@@

Our goal is to reduce the square matrix $A$ by one dimension.
In order to do that, we divide the matrix into four blocks:

@@

thus having two states. Our new two states consist of the original states in the following way:

@@

\\

@@

\\
We can visulize the resulting automata:
    \node[state] (q_0)   {$\mathbf{1}$}; 
    \node[state](q_1) [right=of q_0] {$\mathbf{2}$};
      \path[->] 
      (q_0) edge[bend left, above]  node {A(1,2)} (q_1)
            edge [loop above] node {A(1,1)} ()
      (q_1) edge[bend left, below]  node {A(2,1)} (q_0)
            edge [loop above] node {A(2,2)} ();
Now, instead of calculating $A^*$ with algebra, we can also just look
at the automata and ponder what all possible paths are for $1 \to 1$,
$1 \to 2$, $2 \to 1$, and $2 \to 2$. Note that the automaton is symmetrical, consequently we just need to find paths for $1 \to 1$ and $1 \to 2$.
Starting with $1 \to 1$, we want to find all possible paths from state
one to one with an arbitrary length. Starting at state one, we can either
directly loop back to one, or we transition to two, then loop in two 
arbitrary amount of times and then transition back to one. This whole
process can also be repeated an arbitrary amount of times, yielding the
following formula:
  A(1,1)^* = (A(1,1) + A(1,2)A(2,2)^*A(2,1))^*
Now, because of symmetry:
  A(2,2)^* = (A(2,2) + A(2,1)A(1,1)^*A(1,2))^*
For $1 \to 2$, we can loop in one and then transition to two, in two 
We can either loop there for an arbitrary amount of time or transition to one loop in one an arbitrary amount of times and then go back to
two. This process can also be repeated an arbitrary amount of times.
  A(1,2)^* = A(1,1)^*A(1,2)(A(2,2)^* + A(2,1)A(1,1)^*A(1,2))^*
Again because of symmetry:
  A(2,1)^* = A(2,2)^*A(2,1)(A(1,1)^* + A(1,2)A(2,2)^*A(2,1))^*
The transition matrix $A^*$ has only rational entries, if all blocks
matrices only have rational entries. The block matrices consist of, 
just entries of $A$ that are rational, namely $A(1,1)$, $A(2,2)$, $A(1,2)$
and $A(2,1)$. As well as square matrices, which have
a lower dimension than $n+1$, thus the kleene stars of those are also 
rational because of the induction assumption, namely 
$A(2,2)^*$, $A(1,1)^*$, $(A(1,1)^* + A(1,2)A(2,2)^*A(2,1))^*$, 
$(A(2,2)^* + A(2,1)A(1,1)^*A(1,2))^*$, $(A(2,2) + A(2,1)A(1,1)^*A(1,2))^*$
and $(A(1,1) + A(1,2)A(2,2)^*A(2,1))^*$. Thus, $A^*$ has only rational
entries, which finishes the proof.
\end{proof}\\
\\
Now, we have to prove the other direction:
    {\bf claim: Every rational language can be generated 
    by an finite automata.}
    Previously in this chapter, we defined rational language inductively,  naturally we want to prove the claim with structural induction. That  means we need to construct an automata generating...
    \item ... a language with one letter.
    \item ... the product of two languages.
    \item ... the sum of two languages.
    \item ... the kleene star of one language.
  In the following, all constructed automata will have the first
  state as initial state and all states as accepting state.
  Thus, the sum of the first row of the transition
  matrix $A^*$ will be the accepted language.
  We begin to prove that {\bf (1)} there are an automaton accepting a
  single letter.\\
  Let $\Sigma$ be an arbitrary but fixed alphabet and 
  $\alpha \in \Sigma$ a letter, then:
      \epsilon & \alpha\\
      0        & \epsilon
    Looking at the automata and ponder again, like in the previous proof, we see that $A = A^*$.
      \node[state] (q_0)   {$1$}; 
      \node[state](q_1) [right=of q_0] {$2$};
        \path[->] 
        (q_0) edge[bend left, above]  node {$\alpha$} (q_1)
              edge [loop above] node {$\epsilon$} ()
        (q_1) edge [loop above] node {$\epsilon$} ();
  With our convention,
  we get the accepted language $L = \epsilon + \alpha$, thus $\alpha \in L$.\\
  Next, we will look at the sum of two rational language {\bf (2)}. Let $A,B \in \fps{B}^{n \times n}$ transition matrices of rational languages, then
    B = 
      \epsilon & A_1 & B_1 \\
      0        & A   & 0 \\
      0        & 0   & B
    A_1 = (a_{11}, a_{12}, a_{13}, \dots a_{1n})
    \hspace{0.2cm} \text{and} \hspace{0.2cm}
    B_1 = (b_{11}, b_{12}, b_{13}, \dots b_{1n}).
  \\
  \\
  Depicting the automata: 
      \node[state] (q_0)   {$1$}; 
      \node[state](q_1) [right=of q_0] {$2$};
      \node[state](q_2) [below=of q_0] {$3$};
        \path[->] 
        (q_0) edge[bend left, above]  node {$A_1$} (q_1)
              edge [loop above] node {$\epsilon$} ()
              edge [] node {$B_1$} (q_2)
        (q_1) edge [loop above] node {$A$} ()
        (q_2) edge [loop right] node {$B$}();

    \end{tikzpicture}
  \end{center}
  With the same reasoning as above we get:
  \[
    B^* = 
    \begin{pmatrix}
      \epsilon & A_1A^* & B_1B^*\\
      0        & A^*    & 0 \\
      0        & 0      & B^*
    \end{pmatrix}
  \]
  where 
  \[
    A_1A^* = (\sum_{k=1}^{n}a_{1k}a^*_{k1}, \sum_{k=1}^{n}a_{1k}a^*_{k2},
        \sum_{k=1}^{n}a_{1k}a^*_{k3}, \dots \sum_{k=1}^{n}a_{1k}a^*_{kn})
        = A\hat{}_1
  \]
  and 
  \[
    B_1B^* = (\sum_{k=1}^{n}b_{1k}b^*_{k1}, \sum_{k=1}^{n}b_{1k}b^*_{k2},
        \sum_{k=1}^{n}b_{1k}b^*_{k3}, \dots \sum_{k=1}^{n}b_{1k}b^*_{kn})
        = B{\mathchar"5E}_1.
  \]
  Yielding the first row sum $L = \epsilon + A\hat{}_1 + B{\mathchar"5E}_1$,
  which is exactly the sum of the accepted language of $A$ and $B$.\\
  Next, we will look at the product of two rational language {\bf (3)}.
  Let $p,q\in \mathbb{N}$, $A \in \fps{B}^{p \times p}$ and $B \in \fps{B}^{q \times q}$ transition matrices of rational
  languages, then
  \[
    C = 
    \begin{pmatrix}
      A & PQ \\
      0 & B
    \end{pmatrix}
  \]
  where
  \[ P \in \fps{B}^{p \times q}, \hspace{0.2cm}Q \in \fps{B}^{q \times p}\]
  \[
    P_{ij} = 
    \begin{cases}
      \epsilon & \text{if } \hspace{0.2cm} j = 1\\
      0        & \text{otherwise}
    \end{cases}
    , \hspace{1cm}
    Q_{ij} = 
    \begin{cases}
      \epsilon & \text{if } \hspace{0.2cm} j = 1 \land i = 1\\
      0        & \text{otherwise}
    \end{cases}
  \]
  \\
  \pagebreak
  \\
  Using the formula of the last proof we get:
  \[
    C^* = 
    \begin{pmatrix}
      A^* & A^*PQB^* \\
      0 & B^*
    \end{pmatrix}
  \]
  where
  \[
    (A^*P)_{ij} = 
    \begin{cases}
      \sum_{k=1}^{p} a^*_{1k}p_{k1} =  \sum_{k=1}^{p} a^*_{1k} 
        & \text{if } \hspace{0.2cm} i = 1 \land j = 1\\
      \sum_{k=1}^{p} a^*_{ik}p_{k1} =  \sum_{k=1}^{p} a^*_{ik} 
        & \text{if } \hspace{0.2cm} i > 1\land j = 1\\
      0 & \text{otherwise}
    \end{cases}
  \]
  and 
  \[
    (QB^*)_{ij} = 
    \begin{cases}
      \sum_{k=1}^{p} q_{1k}b^*_{kj} = b^*_{1j}
        & \text{if } \hspace{0.2cm} i = 1\\
      0  & \text{otherwise}
    \end{cases}
  \]
  We are only interested in the first row:
  \[
    (A^*PQB^*)_{1j} = \sum_{k=1}^{q} (A^*P)_{1k}(QB^*)_{kj} 
      = (A^*P)_{11}(QB^*)_{1j}
      = (\sum_{k=1}^{p} a^*_{1k})b^*_{1j}
  \]
  Summing the whole row of $C^*$ together to get the accepted language:

  \[
    L = (\sum_{k=1}^{p} a^*_{1k}) + \sum_{j=1}^q (\sum_{k=1}^{p} a^*_{1k})b^*_{1j}
      = (\sum_{k=1}^{p} a^*_{1k}) + (\sum_{k=1}^{p} a^*_{1k}) (\sum_{j=1}^q b^*_{1j})
      = (\sum_{k=1}^{p} (A^*_1)_k) + (\sum_{k=1}^{p} (A^*_1)_k) (\sum_{j=1}^q (B^*_1)_j)
  \]
  We get the accepted language of $A$ times the accepted language of $B$.\\
  Last but not least, we need to prove that there exists an automaton which
  gives us the kleene star of a rational language. Let 
  $p,q\in \mathbb{N}$ and $A \in \fps{B}^{p \times p}$
  a transition matrix of a rational language 
  and  $P \in \fps{B}^{p \times q}$, $Q \in \fps{B}^{q \times p}$
  like above, then
  \[
    D := 
    \begin{pmatrix}
      E_p &  E_p\\
      0        & A + PQ
    \end{pmatrix}
  \]
  Again, with the formula of the former proof, we get:
  \[
    D^* = 
    \begin{pmatrix}
      E_p &  (A + PQ)^*\\
      0   &  (A + PQ)^*
    \end{pmatrix}
  \]
  The element in the upper right corner can be rewritten:
  \[
    (A+PQ)^* = A^*(PQA^*)^*
  \]
  The sum of the elements in the first row yields $L^*$.
\end{proof}
\section{Conclusion}
The former prove can be modified to also hold for formal power series in
a semi-ring $\fps{S}$ (ref. [\cite{Algebra}]). 
This theorem is called kleene-schützberg theorem (ref. [\cite{Algebra}]) and
it makes a connection between formal power series and weighted automata. 
In summary, formal power series are really powerful, but also quite
unintuitive at first. For futher reading my recommendation is the master thesis
of Laura Wirth (ref. [\cite{master}]).
\pagebreak
\addcontentsline{toc}{section}{Bibliography}
\pagestyle{fancy}
\bibliographystyle{language-dt} 
\bibliography{bibliography} 
%\nocite{*}
\end{document}
